---
title: "Supplemental tables and figures"
output: pdf_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(kableExtra)
library(ggpubr)
load("../RData/publish2.2_actual-summary.RData")
load("../RData/publish2.2_figsS1-S13.RData")
```

Contact: Lia Ossanna, lossanna@arizona.edu or lossanna@nmsu.edu  
Last updated: 2026-03-16  

This document includes all the supplemental tables and figures that accompany "Precipitation moderates associations between native shrubs and invasive buffelgrass (*Pennisetum ciliare*) growth in the Sonoran Desert while local topography has less influence" by Ossanna et al. (2026).


# Tables
## Table S1. Actual values averaged by year
Actual values of variables measured, averaged across all plots for each year (to contextualize year-to-year change values used in analyses). Precipitation measured in mm, buffelgrass density measured in individuals per m^2^, and cover measured in percent.
```{r Table S1}
actual.summary.year2 %>% 
  kbl(booktabs = TRUE, align = "c",
      col.names = c("Year", "Precipitation", "Total culms", "Repro culms", "Density", "Cover", "Shrub cover", "Herb cover")) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>% 
  add_header_above(c(" " = 2, "Buffelgrass" = 4, "Native vegetation" = 2))
```


\newpage
## Table S2. Total culms model 
Model coefficients of the buffelgrass culm count model (response variable: year-to-year change in total number of culms). Note that the reference level for Aspect is east (E). Marginal R^2^ = 0.128; conditional R^2^ = 0.416.
```{r Table S2}
total_avg <- read.table("total_revision1.2.txt", sep = "\t")
colnames(total_avg) <- total_avg[1,]
total_avg <- total_avg[-1,]
row.names(total_avg) <- NULL

total_avg %>% 
  kbl(booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position")) 
```


\newpage
## Table S3. Reproductive culms model 
Model coefficients of the buffelgrass reproductive culm count model (response variable: year-to-year change in number of reproductive culms). Note that the reference level for Aspect is east (E). Marginal R^2^ = 0.100; conditional R^2^ = 0.305.
```{r Table S3}
repro_avg <- read.table("repro_revision1.2.txt", sep = "\t")
colnames(repro_avg) <- repro_avg[1,]
repro_avg <- repro_avg[-1,]
row.names(repro_avg) <- NULL

repro_avg %>% 
  kbl(booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position")) 
```


## Table S4. Density model
Model coefficients of the buffelgrass density model (response variable: year-to-year change in density [individuals per m^2^]). Note that the reference level for Aspect is east (E). Marginal R^2^ = 0.422; conditional R^2^ = 0.468.
```{r Table S4}
bgden_avg <- read.table("bgden_revision1.2.txt", sep = "\t")
colnames(bgden_avg) <- bgden_avg[1,]
bgden_avg <- bgden_avg[-1,]
row.names(bgden_avg) <- NULL

bgden_avg %>% 
  kbl(booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position")) 
```


\newpage
## Table S5. Cover model
Model coefficients of the buffelgrass cover model (response variable: year-to-year-change in cover [%]). Note that the reference level for Aspect is east (E). Marginal R^2^ = 0.261; conditional R^2^ = 0.285.
```{r Table S5}
bgcov_avg <- read.table("bgcov_revision1.2.txt", sep = "\t")
colnames(bgcov_avg) <- bgcov_avg[1,]
bgcov_avg <- bgcov_avg[-1,]
row.names(bgcov_avg) <- NULL

bgcov_avg %>% 
  kbl(booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position")) 
```


## Table S6. Seedling survival model
Model coefficients of the buffelgrass seedling survival model (response variable: seedling survival [%]). Note that the reference level for Aspect is east (E). Marginal R^2^ = 0.493; conditional R^2^ = 0.966.
```{r Table S6}
survival_avg <- read.table("survival_revision1.2.txt", sep = "\t")
colnames(survival_avg) <- survival_avg[1,]
survival_avg <- survival_avg[-1,]
row.names(survival_avg) <- NULL

survival_avg %>% 
  kbl(booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position")) 
```


# Figures
## Figure S1
```{r Fig S1}
precip.site.all
```
**Figure S1.** Precipitation conditions at each site during the study. Red dashed lines indicate mean annual precipitation of each site.

## Figure S2
```{r Fig S2}
repro.shrub.precip
```
**Figure S2.** Relationship between year-to-year change in native shrub cover and year-to-year change in buffelgrass reproductive culm count, moderated by precipitation. Colored lines represent model-predicted values under the wettest precipitation conditions experienced (purple line), the driest conditions (yellow line), and average conditions (green line). Vertical red dashed line indicates no change in shrub cover while horizontal line indicates no change in the number of reproductive buffelgrass culms.


## Figure S3
```{r Fig S3}
repro.herb.precip
```
**Figure S3.** Relationship between year-to-year change in native herbaceous (grass and forb) cover and year-to-year change in buffelgrass reproductive culm count, moderated by precipitation. Colored lines represent model-predicted values under the wettest precipitation conditions experienced (purple line), the driest conditions (yellow line), and average conditions (green line). Vertical red dashed line indicates no change in herb cover while horizontal line indicates no change in the number of reproductive buffelgrass culms. 


## Figure S4
```{r Fig S4, message=FALSE}
ggarrange(shrub.change.precip, herb.change.precip,
          ncol = 2, nrow = 1,
          labels = c("a", "b"))
```
**Figure S4.** Simple linear regression of (A) year-to-year change in native shrub cover and (B) year-to-year change in native herbaceous cover as a function of the previous year’s precipitation, pooling all sampling locations and times. Horizontal red dashed lines indicate no change in shrub or herbaceous cover.


## Figure S5
```{r Fig S5}
total.inbgden
```
**Figure S5.** Relationship between initial buffelgrass density of the plot and year-to-year change in total buffelgrass culm count. Purple line represents model-predicted values. Horizontal red dashed line indicates no change in the number of buffelgrass culms. 


## Figure S6
```{r Fig S6}
repro.inbgden
```
**Figure S6.** Relationship between initial buffelgrass density of the plot and year-to-year change in buffelgrass reproductive culm count. Purple line represents model-predicted values. Horizontal red dashed line indicates no change in the number of reproductive buffelgrass culms. 


## Figure S7
```{r Fig S7}
repro.aspect
```
**Fig. S7.** Effects of aspect on year-to-year change in buffelgrass reproductive culm count. Purple squares represent model-predicted averages. Horizontal red dashed line indicates no change in the number of reproductive buffelgrass culms. 


## Figure S8
```{r Fig S8}
bgden.aspect
```
**Fig. S8.** Effects of aspect on year-to-year change in buffelgrass density. Purple squares represent model-predicted averages. Horizontal red dashed line indicates no change in buffelgrass density.

## Figure S9
```{r Fig S9}
bgcov.aspect
```
**Fig. S9.** Effects of aspect on year-to-year change in buffelgrass cover. Purple squares represent model-predicted averages. Horizontal red dashed line indicates no change in buffelgrass cover.


## Figure S10
```{r Fig S10, message=FALSE}
density.cover
```
**Figure S10.** Simple linear regression of year-to-year change in buffelgrass plot density as a function of year-to-year change in buffelgrass plot cover, pooling all sampling locations and times. Vertical red dashed line indicates no change in buffelgrass cover while horizontal line indicates no change in buffelgrass density.


## Figure S11
```{r Fig S11, message=FALSE}
total.cover
```
**Figure S11.** Simple linear regression of year-to-year change in buffelgrass total culm count as a function of year-to-year change in buffelgrass plot cover, pooling all sampling locations and times. Vertical red dashed line indicates no change in buffelgrass cover while horizontal line indicates no change in buffelgrass total culm count.


## Figure S12
```{r Fig S12}
survival.bgden
```
**Figure S12.** Relationship between buffelgrass seedling survival and buffelgrass density in the plot. Purple line represents model-predicted values.


## Figure S13
```{r Fig S13}
survival.shrub
```
**Figure S13.** Relationship between buffelgrass seedling survival and native shrub cover in the plot. Purple line represents model-predicted values.
