---
title: "Code for linear models"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(ggplot2)
library(lme4)
library(lmerTest)
library(DHARMa)
library(performance)
library(modelbased)
load("../RData/publish2.3_linear-models.RData")
```

Contact: Lia Ossanna, lossanna@arizona.edu or lossanna@nmsu.edu  
Last updated: 2026-05-05

This document includes the R code needed to produce the linear models presented in "Precipitation and slope angle moderate associations between native shrubs and invasive buffelgrass (*Pennisetum ciliare*) growth in the Sonoran Desert" by Ossanna et al. (2026). Code was run with R v4.5.3.

Models:

1. Change in total buffelgrass culm count  
2. Change in reproductive buffelgrass culm count  
3. Change in buffelgrass plot density  
4. Change in buffelgrass plot cover  
5. Buffelgrass seedling survival
6. Simple linear regression to accompany supplemental figures


# Setup
```{r Setup, eval=FALSE}
library(readr) # v2.2.0
library(dplyr) # v1.2.0
library(ggplot2) # v4.0.2
library(lme4) # v1.1-38
library(lmerTest) # v3.2-0
library(DHARMa) # v0.4.7
library(performance) # v0.16.0
library(modelbased) # v0.13.0

# Load data ---------------------------------------------------------------

culm.change.raw <- read_csv("culm-data.csv")
plot.change.raw <- read_csv("plot-data.csv")
survival.dat.raw <- read_csv("survival-data.csv")

# Data wrangling ----------------------------------------------------------

# Center and scale numeric variables for culm change data
culm.change <- culm.change.raw %>% 
  mutate(PlotSlope_scaled = scale(PlotSlope, center = TRUE, scale = TRUE)[, 1],
         Prev_year_precip_scaled = scale(Prev_year_precip, center = TRUE, scale = TRUE)[, 1],
         Change_BGDensity_scaled = scale(Change_BGDensity, center = TRUE, scale = TRUE)[, 1],
         Change_ShrubCover_scaled = scale(Change_ShrubCover, center = TRUE, scale = TRUE)[, 1],
         Change_HerbCover_scaled = scale(Change_HerbCover, center = TRUE, scale = TRUE)[, 1],
         Change_BGCover_scaled = scale(Change_BGCover, center = TRUE, scale = TRUE)[, 1],
         Init_BGDensity_scaled = scale(Init_BGDensity, center = TRUE, scale = TRUE)[, 1],
         Init_ShrubCover_scaled = scale(Init_ShrubCover, center = TRUE, scale = TRUE)[, 1],
         Init_HerbCover_scaled = scale(Init_HerbCover, center = TRUE, scale = TRUE)[, 1],
         Init_BGCover_scaled = scale(Init_BGCover, center = TRUE, scale = TRUE)[, 1])

# Center and scale numeric variables for plot-level data
plot.change <- plot.change.raw %>% 
  mutate(PlotSlope_scaled = scale(PlotSlope, center = TRUE, scale = TRUE)[, 1],
         Prev_year_precip_scaled = scale(Prev_year_precip, center = TRUE, scale = TRUE)[, 1],
         Change_ShrubCover_scaled = scale(Change_ShrubCover, center = TRUE, scale = TRUE)[, 1],
         Change_HerbCover_scaled = scale(Change_HerbCover, center = TRUE, scale = TRUE)[, 1],
         Change_BGCover_scaled = scale(Change_BGCover, center = TRUE, scale = TRUE)[, 1],
         Init_BGDensity_scaled = scale(Init_BGDensity, center = TRUE, scale = TRUE)[, 1],
         Init_ShrubCover_scaled = scale(Init_ShrubCover, center = TRUE, scale = TRUE)[, 1],
         Init_HerbCover_scaled = scale(Init_HerbCover, center = TRUE, scale = TRUE)[, 1],
         Init_BGCover_scaled = scale(Init_BGCover, center = TRUE, scale = TRUE)[, 1])


# Center and scale numeric variables for survival data
dat.survival <- survival.dat.raw %>% 
  mutate(Prev_year_precip_scaled = scale(Prev_year_precip, center = TRUE, scale = TRUE)[, 1],
         PlotSlope_scaled = scale(PlotSlope, center = TRUE, scale = TRUE)[, 1],
         ShrubCover_scaled = scale(ShrubCover, center = TRUE, scale = TRUE)[, 1],
         HerbCover_scaled = scale(HerbCover, center = TRUE, scale = TRUE)[, 1],
         PlotSlope_scaled = scale(PlotSlope, center = TRUE, scale = TRUE)[, 1],
         BGDensity_scaled = scale(BGDensity, center = TRUE, scale = TRUE)[, 1])
```


# 1. Change in total culm count
## Construct model 
```{r total, eval=FALSE}
total <- lmer(Change_TotalCulms ~ Prev_year_precip_scaled +  
                 PlotSlope_scaled + Change_BGDensity_scaled +
                 Change_ShrubCover_scaled + Change_HerbCover_scaled +
                 Prev_year_precip_scaled * Change_BGDensity_scaled +
                 Prev_year_precip_scaled * Change_ShrubCover_scaled +
                 Prev_year_precip_scaled * Change_HerbCover_scaled +
                 Init_BGDensity_scaled + Init_ShrubCover_scaled + Init_HerbCover_scaled +
                 Change_ShrubCover_scaled * PlotSlope_scaled +
                 Change_HerbCover_scaled * PlotSlope_scaled +
                 (1 | Site / Transect / Plot / Plant_ID),
               data = culm.change)
```

## Examine diagnostics
```{r total diagnostics}
summary(total)
r2(total)
res.total <- simulateResiduals(total)
plotQQunif(res.total)
plotResiduals(res.total) 
```


## Predicted vs. observed graph
```{r total pred}
total.pred <- estimate_expectation(total)
total.pred$Change_TotalCulms <- culm.change$Change_TotalCulms

total.pred.plot <- total.pred %>% 
  ggplot(aes(x = Change_TotalCulms, y = Predicted)) +
  geom_point(alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, color = "blue", linewidth = 1) +
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "red") +
  geom_vline(xintercept = 0,
             linetype = "dashed",
             color = "red") +
  ggtitle("Total culm change model, predicted vs. observed") +
  xlab(expression(Delta ~ "Total live culms [observed]")) +
  theme_bw()
total.pred.plot
```



# 2. Change in reproductive culm count
## Construct model
```{r repro, eval=FALSE}
repro <- lmer(Change_ReproductiveCulms ~ Prev_year_precip_scaled +  
                 PlotSlope_scaled + Change_BGDensity_scaled +
                 Change_ShrubCover_scaled + Change_HerbCover_scaled +
                 Prev_year_precip_scaled * Change_BGDensity_scaled +
                 Prev_year_precip_scaled * Change_ShrubCover_scaled +
                 Prev_year_precip_scaled * Change_HerbCover_scaled +
                 Init_BGDensity_scaled + Init_ShrubCover_scaled + Init_HerbCover_scaled +
                 Change_ShrubCover_scaled * PlotSlope_scaled +
                 Change_HerbCover_scaled * PlotSlope_scaled +
                 (1 | Site / Transect / Plot),
               data = culm.change)

```

## Examine diagnostics
```{r repro diagnostics}
summary(repro) 
r2(repro)
res.repro <- simulateResiduals(repro)
plotQQunif(res.repro)
plotResiduals(res.repro) 
```

## Predicted vs. observed graph
```{r repro pred}
repro.pred <- estimate_expectation(repro)
repro.pred$Change_ReproductiveCulms <- culm.change$Change_ReproductiveCulms

repro.pred.plot <- repro.pred %>% 
  ggplot(aes(x = Change_ReproductiveCulms, y = Predicted)) +
  geom_point(alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, color = "blue", linewidth = 1) +
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "red") +
  geom_vline(xintercept = 0,
             linetype = "dashed",
             color = "red") +
  ggtitle("Reproductive culm change model, predicted vs. observed") +
  xlab(expression(Delta ~ "Reproductive culms [observed]")) +
  theme_bw()
repro.pred.plot
```



# 3. Change in plot density
## Construct model
```{r bgden, eval=FALSE}
bgden <- lmer(Change_BGDensity ~ Prev_year_precip_scaled + 
                 PlotSlope_scaled + Change_ShrubCover_scaled +
                 Change_HerbCover_scaled + 
                 Prev_year_precip_scaled * Change_ShrubCover_scaled +
                 Prev_year_precip_scaled * Change_HerbCover_scaled +
                 Init_BGDensity_scaled + Init_ShrubCover_scaled + Init_HerbCover_scaled +
                 Change_ShrubCover_scaled * PlotSlope_scaled +
                 Change_HerbCover_scaled * PlotSlope_scaled +
                 (1 | Site / Transect),
               data = plot.change)
```

## Examine dianostics
```{r bgden diagnostics}
summary(bgden)
r2(bgden)
res.bgden <- simulateResiduals(bgden)
plotQQunif(res.bgden)
plotResiduals(res.bgden) 
check_collinearity(bgden)
```

## Predicted vs. observed graph
```{r bgden pred}
bgden.pred <- estimate_expectation(bgden)
bgden.pred$Change_BGDensity <- plot.change$Change_BGDensity

bgden.pred.plot <- bgden.pred %>% 
  ggplot(aes(x = Change_BGDensity, y = Predicted)) +
  geom_point(alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, color = "blue", linewidth = 1) +
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "red") +
  geom_vline(xintercept = 0,
             linetype = "dashed",
             color = "red") +
  ggtitle("Density change model, predicted vs. observed") +
  xlab(expression(Delta ~ paste("Buffelgrass density (individuals / ", m^2, ") [observed]"))) +
  theme_bw()
bgden.pred.plot
```



# 4. Change in plot cover
## Construct model
```{r bgcov, eval=FALSE}
bgcov <- lmer(Change_BGCover ~ Prev_year_precip_scaled + 
                 PlotSlope_scaled + Change_ShrubCover_scaled +
                 Change_HerbCover_scaled + 
                 Prev_year_precip_scaled * Change_ShrubCover_scaled +
                 Prev_year_precip_scaled * Change_HerbCover_scaled +
                 Init_BGCover_scaled + Init_ShrubCover_scaled + Init_HerbCover_scaled +
                 Change_ShrubCover_scaled * PlotSlope_scaled +
                 Change_HerbCover_scaled * PlotSlope_scaled +
                 (1 | Site),
               data = plot.change)
```

## Examine diagnostics
```{r bgcov diagnostics}
summary(bgcov)
r2(bgcov)
res.bgcov <- simulateResiduals(bgcov)
plotQQunif(res.bgcov)
plotResiduals(res.bgcov) 
```

## Predicted vs. observed graph
```{r bgcov pred}
bgcov.pred <- estimate_expectation(bgcov)
bgcov.pred$Change_BGCover <- plot.change$Change_BGCover

bgcov.pred.plot <- bgcov.pred %>% 
  ggplot(aes(x = Change_BGCover, y = Predicted)) +
  geom_point(alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, color = "blue", linewidth = 1) +
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "red") +
  geom_vline(xintercept = 0,
             linetype = "dashed",
             color = "red") +
  ggtitle("Cover change model, predicted vs. observed") +
  xlab(expression(Delta ~ "Buffelgrass cover (%) [observed]")) +
  theme_bw()
bgcov.pred.plot
```



# 5. Seedling survival
## Construct model
```{r survival, eval=FALSE}
survival <- glmer(cbind(seedlings_surviving,
                         remaining_toothpicks - seedlings_surviving) ~ Prev_year_precip_scaled + 
                     PlotSlope_scaled + BGDensity_scaled +
                     ShrubCover_scaled + HerbCover_scaled + 
                     Prev_year_precip_scaled * BGDensity_scaled + 
                     Prev_year_precip_scaled * ShrubCover_scaled +
                     Prev_year_precip_scaled * HerbCover_scaled +
                     ShrubCover_scaled * PlotSlope_scaled +
                     HerbCover_scaled * PlotSlope_scaled +
                     (1 | Site / Transect / Plot),
                   family = binomial,
                   data = dat.survival)
```

## Examine diagnostics
```{r survival diagnostics}
summary(survival)
r2(survival) 
res.survival <- simulateResiduals(survival)
plotQQunif(res.survival)
plotResiduals(res.survival) 
check_collinearity(survival)
```

## Predicted vs. observed graph
```{r survival pred}
survival.pred <- estimate_expectation(survival)
survival.pred$survival_prop <- dat.survival$survival_prop

survival.pred.plot <- survival.pred %>% 
  ggplot(aes(x = survival_prop, y = Predicted)) +
  geom_point(alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, color = "blue", linewidth = 1) +
  ggtitle("Seedling survival model, predicted vs. observed")  +
  xlab("Proportion of surviving seedlings [observed]") +
  theme_bw() 
survival.pred.plot
```

# 6. Simple linear regression
## Shrub cover change vs. precipitation
To accompany Figure S3a.
```{r}
summary(lm(Change_ShrubCover ~ Prev_year_precip, data = plot.change))
```

## Grass & forb cover change vs. precipitation
To accompany Figure S3b.
```{r}
summary(lm(Change_HerbCover ~ Prev_year_precip, data = plot.change))
```

## Buffelgrass density change vs. cover change
To accompany Figure S8.
```{r}
summary(lm(Change_BGDensity ~ Change_BGCover, data = culm.change))
```

## Buffelgrass culm count change vs. cover change
To accompany Figure S9.
```{r}
summary(lm(Change_TotalCulms ~ Change_BGCover, data = culm.change))
```

