---
title: "Supplemental tables and figures"
output: pdf_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(kableExtra)
load("../RData/publish_model-averages.RData")
load("../RData/publish_actual-summary.RData")
```

Includes all supplemental tables and figures to accompany "Precipitation moderates the effect of native shrubs on invasive buffelgrass (*Pennisetum ciliare*) growth in the Sonoran Desert while local topography has less influence" by Ossanna et al. (2026).

# **Table S1**. Actual values
Actual values of variables measured, averaged across all plots for each year (to contextualize year-to-year change values used in analyses). Precipitation measured in mm, buffelgrass density measured in individuals per m^2^, and cover measured in percent.
```{r Table S1}
actual.summary.year2 %>% 
  kbl(booktabs = TRUE, align = "c",
      col.names = c("Year", "Precipitation", "Total culms", "Repro culms", "Density", "Cover", "Shrub cover", "Herb cover")) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>% 
  add_header_above(c(" " = 2, "Buffelgrass" = 4, "Native vegetation" = 2))
```


# **Table S2**. Total culms model
Model-averaged coefficients of the buffelgrass culm count model calculated from two top models (response variable: year-to-year change in total number of culms). Marginal R^2^ = 0.144–0.147; conditional R^2^
= 0.386–0.391.
```{r Table S2}
summary(total_avg)
```


