---
title: "Code for supplemental figures"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(ggplot2)
library(insight)
library(viridis)
library(ggpubr)
load("../RData/publish2.3_supp-figures.RData")
```

Contact: Lia Ossanna, lossanna@arizona.edu or lossanna@nmsu.edu  
Last updated: 2026-05-05

This document includes the R code needed to produce the supplemental figures presented in "Precipitation and slope angle moderate associations between native shrubs and invasive buffelgrass (*Pennisetum ciliare*) growth in the Sonoran Desert" by Ossanna et al. (2026). Code was run with R v4.5.3.

For detailed information and annotated code, see `code-for-main-figures.html`.

Figures:

- S1: Precipitation conditions
- S2: Reproductive culm change vs. herb cover change
- S3: Shrub/herb cover change vs. precip
- S4: Total culm change vs. density change
- S5: Total culm change vs. initial density
- S6: Reproductive culm change vs. initial density
- S7: Cover change vs. initial cover
- S8: Density change vs. cover change
- S9: Total culm change vs. cover change
- S10: Survival vs. density
- S11: Survival vs. shrub cover

# Setup
```{r Setup, eval=FALSE}
library(readr) # v2.2.0
library(dplyr) # v1.2.0
library(ggplot2) # v4.0.2
library(insight) # v1.3.1
library(viridis) # v0.6.5
library(ggpubr) # v0.6.3

# Load data ---------------------------------------------------------------

culm.change.raw <- read_csv("culm-data.csv")
plot.change.raw <- read_csv("plot-data.csv")
survival.dat.raw <- read_csv("survival-data.csv")
dat <- read_csv("all-data.csv")

# Center and scale numeric variables --------------------------------------

# Center and scale numeric variables for culm change data
culm.change <- culm.change.raw %>% 
  mutate(PlotSlope_scaled = scale(PlotSlope, center = TRUE, scale = TRUE)[, 1],
         Prev_year_precip_scaled = scale(Prev_year_precip, center = TRUE, scale = TRUE)[, 1],
         Change_BGDensity_scaled = scale(Change_BGDensity, center = TRUE, scale = TRUE)[, 1],
         Change_ShrubCover_scaled = scale(Change_ShrubCover, center = TRUE, scale = TRUE)[, 1],
         Change_HerbCover_scaled = scale(Change_HerbCover, center = TRUE, scale = TRUE)[, 1],
         Change_BGCover_scaled = scale(Change_BGCover, center = TRUE, scale = TRUE)[, 1],
         Init_BGDensity_scaled = scale(Init_BGDensity, center = TRUE, scale = TRUE)[, 1],
         Init_ShrubCover_scaled = scale(Init_ShrubCover, center = TRUE, scale = TRUE)[, 1],
         Init_HerbCover_scaled = scale(Init_HerbCover, center = TRUE, scale = TRUE)[, 1],
         Init_BGCover_scaled = scale(Init_BGCover, center = TRUE, scale = TRUE)[, 1])

# Center and scale numeric variables for plot-level data
plot.change <- plot.change.raw %>% 
  mutate(PlotSlope_scaled = scale(PlotSlope, center = TRUE, scale = TRUE)[, 1],
         Prev_year_precip_scaled = scale(Prev_year_precip, center = TRUE, scale = TRUE)[, 1],
         Change_ShrubCover_scaled = scale(Change_ShrubCover, center = TRUE, scale = TRUE)[, 1],
         Change_HerbCover_scaled = scale(Change_HerbCover, center = TRUE, scale = TRUE)[, 1],
         Change_BGCover_scaled = scale(Change_BGCover, center = TRUE, scale = TRUE)[, 1],
         Init_BGDensity_scaled = scale(Init_BGDensity, center = TRUE, scale = TRUE)[, 1],
         Init_ShrubCover_scaled = scale(Init_ShrubCover, center = TRUE, scale = TRUE)[, 1],
         Init_HerbCover_scaled = scale(Init_HerbCover, center = TRUE, scale = TRUE)[, 1],
         Init_BGCover_scaled = scale(Init_BGCover, center = TRUE, scale = TRUE)[, 1])


# Center and scale numeric variables for survival data
dat.survival <- survival.dat.raw %>% 
  mutate(Prev_year_precip_scaled = scale(Prev_year_precip, center = TRUE, scale = TRUE)[, 1],
         PlotSlope_scaled = scale(PlotSlope, center = TRUE, scale = TRUE)[, 1],
         ShrubCover_scaled = scale(ShrubCover, center = TRUE, scale = TRUE)[, 1],
         HerbCover_scaled = scale(HerbCover, center = TRUE, scale = TRUE)[, 1],
         PlotSlope_scaled = scale(PlotSlope, center = TRUE, scale = TRUE)[, 1],
         BGDensity_scaled = scale(BGDensity, center = TRUE, scale = TRUE)[, 1])



# Construct 3 dataset versions needed for graphing ------------------------

# Need 3 datasets:
#   1. Data for graphing observed data
#   2. Data to construct datagrid with model predictions
#   3. Data to construct datagrid with unscaled explanatory variable (to be added to #2)


# 1. Dataset for graphing (response and unscaled explanatory variables)
#   Culm change - total & repro
dat.culm <- culm.change %>% 
  select(Change_TotalCulms, Change_ReproductiveCulms, 
         Prev_year_precip, PlotSlope, Change_BGDensity, Change_ShrubCover, Change_HerbCover,
         Init_BGDensity, Init_ShrubCover, Init_HerbCover)

#   Plot change - density & cover
dat.bgden <- plot.change %>% 
  select(Change_BGDensity,  
         Prev_year_precip, PlotSlope, Change_ShrubCover, Change_HerbCover,
         Init_BGDensity, Init_ShrubCover, Init_HerbCover) 

dat.bgcov <- plot.change %>% 
  select(Change_BGCover, 
         Prev_year_precip, PlotSlope, Change_ShrubCover, Change_HerbCover,
         Init_BGCover, Init_ShrubCover, Init_HerbCover) 


#   Survival  
dat.survival <- survival.dat %>%
  select(remaining_toothpicks, seedlings_surviving, survival_prop,
         Prev_year_precip, PlotSlope, BGDensity, ShrubCover, HerbCover,
         Prev_year_precip_scaled, PlotSlope_scaled, BGDensity_scaled,
         ShrubCover_scaled, HerbCover_scaled)


# 2. Dataset for constructing datagrid with predictions (scaled explanatory variables only)
#   Culm change - total & repro 
dat.culm.ex <- culm.change %>% 
  select(Prev_year_precip_scaled, PlotSlope_scaled, 
         Change_BGDensity_scaled, Change_ShrubCover_scaled, Change_HerbCover_scaled,
         Init_BGDensity_scaled, Init_ShrubCover_scaled, Init_HerbCover_scaled)

#   Plot change - density & cover
dat.bgden.ex <- plot.change %>% 
  select(Prev_year_precip_scaled,
         PlotSlope_scaled, Change_ShrubCover_scaled, Change_HerbCover_scaled,
         Init_BGDensity_scaled, Init_ShrubCover_scaled, Init_HerbCover_scaled)

dat.bgcov.ex <- plot.change %>% 
  select(Prev_year_precip_scaled,
         PlotSlope_scaled, Change_ShrubCover_scaled, Change_HerbCover_scaled,
         Init_BGCover_scaled, Init_ShrubCover_scaled, Init_HerbCover_scaled)

#   Survival  
dat.survival.ex <- survival.dat %>%
  select(Prev_year_precip_scaled,
         PlotSlope_scaled, BGDensity_scaled, ShrubCover_scaled, HerbCover_scaled)


# 3. Dataset for constructing datagrid with unscaled variables to match graph (unscaled explanatory variables only)
#   Culm change - total & repro
dat.culm.unscaled <- culm.change %>% 
  select(Prev_year_precip, PlotSlope, 
         Change_BGDensity, Change_ShrubCover, Change_HerbCover,
         Init_BGDensity, Init_ShrubCover, Init_HerbCover)

#   Plot change - density & cover
dat.bgden.unscaled <- plot.change %>% 
  select(Prev_year_precip, PlotSlope, Change_ShrubCover, Change_HerbCover,
         Init_BGDensity, Init_ShrubCover, Init_HerbCover)

dat.bgcov.unscaled <- plot.change %>% 
  select(Prev_year_precip, PlotSlope, Change_ShrubCover, Change_HerbCover,
         Init_BGCover, Init_ShrubCover, Init_HerbCover)

#   Survival 
dat.survival.unscaled <- survival.dat %>%
  select(Prev_year_precip, PlotSlope, BGDensity, ShrubCover, HerbCover)
```


# Figure S1: Precipitation conditions
```{r}
# Create dataframe of dat for with precip cols
dat.precip <- dat %>% 
  select(Year, Site, Transect, Plot, Prev_year_precip, Perc_dev, MAP) %>% 
  distinct(.keep_all = TRUE) 

# Create dataframe of average for precip by site (red dashed lines)
precip.avg.site <- dat.precip %>% 
  group_by(Year, Site) %>% 
  summarise(Perc_dev_avg = mean(Perc_dev),
            Prev_year_precip_avg = mean(Prev_year_precip),
            .groups = "keep")

map.avg <- dat.precip %>% 
  select(Site, Transect, MAP) %>% 
  distinct(.keep_all = TRUE) %>% 
  group_by(Site) %>% 
  summarise(MAP_avg = mean(MAP))

# Line graph of averages across sites
precip.site.all <- precip.avg.site %>% 
  ggplot(aes(x = Year, y = Prev_year_precip_avg)) +
  geom_point() +
  geom_line() +
  facet_wrap(~Site) +
  theme_bw() +
  labs(x = NULL,
       y = "Previous year precip (mm)",
       title = "Precipitation conditions") +
  theme(axis.text.x = element_text(color = "black")) +
  geom_hline(data = map.avg, 
             aes(yintercept = MAP_avg), 
             linetype = "dashed", color = "red")
precip.site.all
```


# Figure S2: Reproductive culm change vs. herb cover change
## Create a dataframe to graph model-predicted lines
```{r, message=FALSE}
# Generate prediction and add unscaled variable
insight.repro.herb <- get_datagrid(dat.culm.ex, by = c("Change_HerbCover_scaled"),
                                   length = 100)
insight.repro.herb$Change_ReproductiveCulms <- get_predicted(repro, insight.repro.herb)
unscaled.herb <- get_datagrid(dat.culm.unscaled, by = "Change_HerbCover",
                              length = 100) %>% 
  arrange(Change_HerbCover)
insight.repro.herb$Change_HerbCover <- unscaled.herb$Change_HerbCover
```

## Graph
```{r}
repro.herb <- dat.culm %>% 
  ggplot(aes(x = Change_HerbCover, y = Change_ReproductiveCulms)) +
  geom_point(alpha = 0.35) +
  geom_line(data = insight.repro.herb,
            aes(y = Change_ReproductiveCulms), linewidth = 1,
            color = "purple3") +
  theme_bw() +
  labs(y = expression(Delta ~ "Buffelgrass reproductive culm count"),
       x = expression(Delta ~ "Native grass & forb cover (%)")) +
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "red") +
  geom_vline(xintercept = 0,
             linetype = "dashed",
             color = "red") 
repro.herb
```


# Figure S3: Shrub/herb cover change vs. precip
## 3a: Shrub
```{r, message=FALSE}
shrub.change.precip <- plot.change %>% 
  ggplot(aes(x = Prev_year_precip, y = Change_ShrubCover)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(y = expression(Delta ~ "Native shrub cover (%)"),
       x = "Previous year precipitation (mm)") +
  theme(plot.margin = margin(10, 10, 10, 20))
```

## 3b: Herb
```{r, message=FALSE}
herb.change.precip <- plot.change %>% 
  ggplot(aes(x = Prev_year_precip, y = Change_HerbCover)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(y = expression(Delta ~ "Native grass & forb cover (%)"),
       x = "Previous year precipitation (mm)") +
  theme(plot.margin = margin(10, 10, 10, 20))
```

## Combine
```{r, message=FALSE}
ggarrange(shrub.change.precip, herb.change.precip,
          ncol = 2, nrow = 1,
          labels = c("a", "b"))
```


# Figure S4: Total culm change vs. density change
## Create a dataframe to graph model-predicted lines
```{r}
# Generate prediction and add unscaled variable
insight.total.bgden <- get_datagrid(dat.culm.ex, by = c("Change_BGDensity_scaled"),
                                    length = 100)
insight.total.bgden$Change_TotalCulms <- get_predicted(total, insight.total.bgden)
unscaled.bgden <- get_datagrid(dat.culm.unscaled, by = "Change_BGDensity",
                               length = 100) %>% 
  arrange(Change_BGDensity)
insight.total.bgden$Change_BGDensity <- unscaled.bgden$Change_BGDensity
```

## Graph
```{r}
total.bgden <- dat.culm %>% 
  ggplot(aes(x = Change_BGDensity, y = Change_TotalCulms)) +
  geom_point(alpha = 0.35) +
  geom_line(data = insight.total.bgden,
            aes(y = Change_TotalCulms), linewidth = 1,
            color = "purple3") +
  theme_bw() +
  labs(y = expression(Delta ~ "Total buffelgrass culm count"),
       x = expression(Delta ~ paste("Buffelgrass density (individuals / ", m^2, ")"))) +
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "red") +
  geom_vline(xintercept = 0,
             linetype = "dashed",
             color = "red") 
total.bgden
```



# Figure S5: Total culm change vs. initial density
## Create a dataframe to graph model-predicted lines
```{r}
# Generate prediction and add unscaled variable
insight.total.inbgden <- get_datagrid(dat.culm.ex, by = c("Init_BGDensity_scaled"),
                                      length = 100)
insight.total.inbgden$Change_TotalCulms <- get_predicted(total, insight.total.inbgden)
unscaled.inbgden <- get_datagrid(dat.culm.unscaled, by = "Init_BGDensity",
                                 length = 100) %>% 
  arrange(Init_BGDensity)
insight.total.inbgden$Init_BGDensity <- unscaled.inbgden$Init_BGDensity
```

## Graph
```{r}
total.inbgden <- dat.culm %>% 
  ggplot(aes(x = Init_BGDensity, y = Change_TotalCulms)) +
  geom_point(alpha = 0.35) +
  geom_line(data = insight.total.inbgden,
            aes(y = Change_TotalCulms), linewidth = 1,
            color = "purple3") +
  theme_bw() +
  labs(x = expression(paste("Initial buffelgrass density (individuals / ", m^2, ")")),
       y = expression(Delta ~ "Total buffelgrass culm count")) +
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "red")
total.inbgden
```



# Figure S6: Reproductive culm change vs. initial density
## Create a dataframe to graph model-predicted lines
```{r}
# Generate prediction and add unscaled variable
insight.repro.inbgden <- get_datagrid(dat.culm.ex, by = c("Init_BGDensity_scaled"),
                                      length = 100)
insight.repro.inbgden$Change_ReproductiveCulms <- get_predicted(repro, insight.repro.inbgden)
unscaled.inbgden <- get_datagrid(dat.culm.unscaled, by = "Init_BGDensity",
                                 length = 100) %>% 
  arrange(Init_BGDensity)
insight.repro.inbgden$Init_BGDensity <- unscaled.inbgden$Init_BGDensity
```

## Graph
```{r}
repro.inbgden <- dat.culm %>% 
  ggplot(aes(x = Init_BGDensity, y = Change_ReproductiveCulms)) +
  geom_point(alpha = 0.35) +
  geom_line(data = insight.repro.inbgden,
            aes(y = Change_ReproductiveCulms), linewidth = 1,
            color = "purple3") +
  theme_bw() +
  labs(x = expression(paste("Initial buffelgrass density (individuals / ", m^2, ")")),
       y = expression(Delta ~ "Buffelgrass reproductive culm count")) +
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "red")
repro.inbgden
```



# Figure S7: Cover change vs. initial cover
## Create a dataframe to graph model-predicted lines
```{r}
# Generate prediction and add unscaled variable 
insight.bgcov.inbgcov <- get_datagrid(dat.bgcov.ex, by = c("Init_BGCover_scaled"),
                                      length = 10)
insight.bgcov.inbgcov$Change_BGCover <- get_predicted(bgcov, insight.bgcov.inbgcov)
unscaled.inbgcov <- get_datagrid(dat.bgcov.unscaled, by = "Init_BGCover",
                                 length = 10) %>% 
  arrange(Init_BGCover)
insight.bgcov.inbgcov$Init_BGCover <- unscaled.inbgcov$Init_BGCover
```

## Graph
```{r}
bgcov.inbgcov <- dat.bgcov %>% 
  ggplot(aes(x = Init_BGCover, y = Change_BGCover)) +
  geom_point(alpha = 0.7) +
  geom_line(data = insight.bgcov.inbgcov,
            aes(y = Change_BGCover), linewidth = 1,
            color = "purple3") +
  theme_bw() +
  labs(y = expression(Delta ~ "Buffelgrass cover (%)"),
       x = expression("Initial buffelgrass cover (%)")) +
  geom_hline(yintercept = 0,
             linetype = "dashed",
             color = "red") 
bgcov.inbgcov
```



# Figure S8: Density change vs. cover change
```{r, message=FALSE}
density.cover <- plot.change %>% 
  ggplot(aes(x = Change_BGCover, y = Change_BGDensity)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") +
  labs(x = expression(Delta ~ "Buffelgrass cover (%)"),
       y = expression(Delta ~ paste("Buffelgrass density (individuals / ", m^2, ")")))
density.cover
```


# Figure S9: Total culm change vs. cover change
```{r, message=FALSE}
total.cover <- culm.change %>% 
  ggplot(aes(x = Change_BGCover, y = Change_TotalCulms)) +
  geom_point(alpha = 0.35) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") +
  labs(y = expression(Delta ~ "Total buffelgrass culm count"),
       x = expression(Delta ~ "Buffelgrass cover (%)"))
total.cover
```



# Figure S10: Survival vs. density
## Create a dataframe to graph model-predicted lines
```{r}
insight.survival.bgden <- get_datagrid(dat.survival.ex, by = c("BGDensity_scaled"),
                                       length = 50)
insight.survival.bgden$survival_prop <- get_predicted(survival, insight.survival.bgden)
unscaled.bgden <- get_datagrid(dat.survival.unscaled, by = "BGDensity",
                               length = 50) %>%
  arrange(BGDensity)
insight.survival.bgden$BGDensity <- unscaled.bgden$BGDensity
```

## Graph
```{r}
survival.bgden <- dat.survival %>%
  ggplot(aes(x = BGDensity, y = survival_prop)) +
  geom_point(alpha = 0.7) +
  geom_line(data = insight.survival.bgden,
            aes(y = survival_prop), linewidth = 1,
            color = "purple3") +
  theme_bw() +
  labs(x = expression(paste("Buffelgrass density (individuals / ", m^2, ")")),
       y = "Proportion of surviving buffelgrass seedlings")
survival.bgden
```


# Figure S11: Survival vs. shrub cover
## Create a dataframe to graph model-predicted lines
```{r}
insight.survival.shrub <- get_datagrid(dat.survival.ex, by = c("ShrubCover_scaled"),
                                       length = 50)
insight.survival.shrub$survival_prop <- get_predicted(survival, insight.survival.shrub) 
unscaled.shrub <- get_datagrid(dat.survival.unscaled, by = "ShrubCover",
                               length = 50) %>%
  arrange(ShrubCover)
insight.survival.shrub$ShrubCover <- unscaled.shrub$ShrubCover
```

## Graph
```{r}
survival.shrub <- dat.survival %>%
  ggplot(aes(x = ShrubCover, y = survival_prop)) +
  geom_point(alpha = 0.7) +
  geom_line(data = insight.survival.shrub,
            aes(y = survival_prop), linewidth = 1,
            color = "purple3") +
  theme_bw() +
  labs(x = "Native shrub cover (%)",
       y = "Proportion of surviving buffelgrass seedlings")
survival.shrub
```

