---
title: "Supplemental tables and figures"
output: pdf_document
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(kableExtra)
library(ggplot2)
library(ggpubr)
load("../RData/publish_actual-summary.RData")
load("../RData/publish_figsS1-S9.RData")
```

Includes all supplemental tables and figures to accompany "Precipitation moderates the effect of native shrubs on invasive buffelgrass (*Pennisetum ciliare*) growth in the Sonoran Desert while local topography has less influence" by Ossanna et al. (2026).

Contact: Lia Ossanna, lossanna@arizona.edu or lossanna@nmsu.edu  
Last updated: 2025-09-23


# Tables
## Table S1. Actual values averaged by year
Actual values of variables measured, averaged across all plots for each year (to contextualize year-to-year change values used in analyses). Precipitation measured in mm, buffelgrass density measured in individuals per m^2^, and cover measured in percent.
```{r Table S1}
actual.summary.year2 %>% 
  kbl(booktabs = TRUE, align = "c",
      col.names = c("Year", "Precipitation", "Total culms", "Repro culms", "Density", "Cover", "Shrub cover", "Herb cover")) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>% 
  add_header_above(c(" " = 2, "Buffelgrass" = 4, "Native vegetation" = 2))
```


## Table S2. Total culms model 
Model-averaged coefficients of the buffelgrass culm count model calculated from two top models (response variable: year-to-year change in total number of culms). Marginal R^2^ = 0.144–0.147; conditional  
R^2^ = 0.386–0.391.
```{r Table S2}
total_avg <- read.table("total_avg.txt", sep = "\t")
colnames(total_avg) <- total_avg[1,]
total_avg <- total_avg[-1,]
row.names(total_avg) <- NULL

total_avg %>% 
  kbl(booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position")) 
```


## Table S3. Reproductive culms model 
Model-averaged coefficients of the buffelgrass reproductive culm count model calculated from five top models (response variable: year-to-year change in number of reproductive culms). Marginal R^2^ = 0.147–0.155; conditional R^2^ = 0.318–0.333.
```{r Table S3}
repro_avg <- read.table("repro_avg.txt", sep = "\t")
colnames(repro_avg) <- repro_avg[1,]
repro_avg <- repro_avg[-1,]
row.names(repro_avg) <- NULL

repro_avg %>% 
  kbl(booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position")) 
```


## Table S4. Plot density model
Model-averaged coefficients of the buffelgrass plot density model calculated from five top models (response variable: year-to-year change in plot density [individuals per m^2^]). Marginal R^2^ = 0.360–0.389; conditional R^2^ = 0.423–0.458.
```{r Table S4}
bgden_avg <- read.table("bgden_avg.txt", sep = "\t")
colnames(bgden_avg) <- bgden_avg[1,]
bgden_avg <- bgden_avg[-1,]
row.names(bgden_avg) <- NULL

bgden_avg %>% 
  kbl(booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position")) 
```


\newpage
## Table S5. Plot cover model
Model-averaged coefficients of the buffelgrass plot cover model calculated from six top models (response variable: year-to-year-change in plot cover [%]). Marginal R^2^ = 0.243–0.246; conditional R^2^ = 0.284–0.289.
```{r Table S5}
bgcov_avg <- read.table("bgcov_avg.txt", sep = "\t")
colnames(bgcov_avg) <- bgcov_avg[1,]
bgcov_avg <- bgcov_avg[-1,]
row.names(bgcov_avg) <- NULL

bgcov_avg %>% 
  kbl(booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position")) 
```


## Table S6. Seedling survival model
Model-averaged coefficients of the buffelgrass seedling survival model calculated from 15 top models (response variable: seedling survival [%]). Marginal R^2^ = 0.666–0.695; conditional R^2^ = 0.779–0.800.
```{r Table S6}
survival_avg <- read.table("survival_avg.txt", sep = "\t")
colnames(survival_avg) <- survival_avg[1,]
survival_avg <- survival_avg[-1,]
row.names(survival_avg) <- NULL

survival_avg %>% 
  kbl(booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position")) 
```


# Figures
## Figure S1
```{r Fig S1}
repro.shrub.precip
```
**Figure S1.** Effects of year-to-year change in native shrub cover on year-to-year change in reproductive buffelgrass culm count, moderated by precipitation. Colored lines represent model-predicted values under the wettest precipitation conditions experienced (purple line), the driest conditions (yellow line), and average conditions (green line). Vertical red dashed line indicates no change in shrub cover while horizontal line indicates no change in the number of reproductive buffelgrass culms.

\newpage
## Figure S2
```{r Fig S2}
repro.herb.precip
```
**Figure S2.** Effects of year-to-year change in native herbaceous (grass and forb) cover on year-to-year change in reproductive buffelgrass culm count, moderated by precipitation. Colored lines represent model-predicted values under the wettest precipitation conditions experienced (purple line), the driest conditions (yellow line), and average conditions (green line). Vertical red dashed line indicates no change in herb cover while horizontal line indicates no change in the number of reproductive buffelgrass culms. 

\newpage
## Figure S3
```{r Fig S3, message=FALSE}
ggarrange(shrub.change.precip, herb.change.precip,
          ncol = 2, nrow = 1,
          labels = c("(A)", "(B)"))
```
**Figure S3.** Simple linear regression of (A) year-to-year change in native shrub cover and (B) year-to-year change in native herbaceous cover as a function of the previous year’s precipitation, pooling all sampling locations and times. Horizontal red dashed lines indicate no change in shrub or herbaceous cover.

\newpage
## Figure S4
```{r Fig S4}
repro.aspect
```
**Fig. S4.** Effects of aspect on year-to-year change in buffelgrass reproductive culm count. Purple squares represent model-predicted averages. Horizontal red dashed line indicates no change in the number of reproductive buffelgrass culms. 

\newpage
## Figure S5
```{r Fig S5}
bgden.aspect
```
**Fig. S5.** Effects of aspect on year-to-year change in buffelgrass plot density. Purple squares represent model-predicted averages. Horizontal red dashed line indicates no change in buffelgrass density.


\newpage
## Figure S6
```{r Fig S6}
bgcov.aspect
```
**Fig. S6.** Effects of aspect on year-to-year change in buffelgrass plot cover. Purple squares represent model-predicted averages. Horizontal red dashed line indicates no change in buffelgrass cover.


\newpage
## Figure S7
```{r Fig S7}
bgcov.shrub.precip
```
**Figure S7.** Effects of year-to-year change in native shrub cover on year-to-year change in buffelgrass plot cover, moderated by precipitation. Colored lines represent model-predicted values under the wettest precipitation conditions experienced (purple line), the driest conditions (yellow line), and average conditions (green line). Vertical red dashed line indicates no change in shrub cover while horizontal line indicates no change in buffelgrass cover.

\newpage
## Figure S8
```{r Fig S8, message=FALSE}
density.cover
```
**Figure S8.** Simple linear regression of year-to-year change in buffelgrass plot density as a function of year-to-year change in buffelgrass plot cover, pooling all sampling locations and times. Vertical red dashed line indicates no change in buffelgrass cover while horizontal line indicates no change in buffelgrass density.

\newpage
## Figure S9
```{r Fig S9, message=FALSE}
total.cover
```
**Figure S9.** Simple linear regression of year-to-year change in buffelgrass total culm count as a function of year-to-year change in buffelgrass plot cover, pooling all sampling locations and times. Vertical red dashed line indicates no change in buffelgrass cover while horizontal line indicates no change in buffelgrass total culm count.
