---
title: "Supplemental tables and figures"
output: pdf_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(kableExtra)
load("../RData/publish_actual-summary.RData")
```

Includes all supplemental tables and figures to accompany "Precipitation moderates the effect of native shrubs on invasive buffelgrass (*Pennisetum ciliare*) growth in the Sonoran Desert while local topography has less influence" by Ossanna et al. (2026).


# Table S1. Actual values averaged by year
Actual values of variables measured, averaged across all plots for each year (to contextualize year-to-year change values used in analyses). Precipitation measured in mm, buffelgrass density measured in individuals per m^2^, and cover measured in percent.
```{r Table S1}
actual.summary.year2 %>% 
  kbl(booktabs = TRUE, align = "c",
      col.names = c("Year", "Precipitation", "Total culms", "Repro culms", "Density", "Cover", "Shrub cover", "Herb cover")) %>%
  kable_styling(latex_options = c("striped", "hold_position")) %>% 
  add_header_above(c(" " = 2, "Buffelgrass" = 4, "Native vegetation" = 2))
```


# Table S2. Total culms model 
Model-averaged coefficients of the buffelgrass culm count model calculated from two top models (response variable: year-to-year change in total number of culms). Marginal R^2^ = 0.144–0.147; conditional  
R^2^ = 0.386–0.391.
```{r Table S2}
total_avg <- read.table("total_avg.txt", sep = "\t")
colnames(total_avg) <- total_avg[1,]
total_avg <- total_avg[-1,]
row.names(total_avg) <- NULL

total_avg %>% 
  kbl(booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position")) 
```


# Table S3. Reproductive culms model 
Model-averaged coefficients of the buffelgrass reproductive culm count model calculated from five top models (response variable: year-to-year change in number of reproductive culms). Marginal R^2^ = 0.147–0.155; conditional R^2^ = 0.318–0.333.
```{r Table S3}
repro_avg <- read.table("repro_avg.txt", sep = "\t")
colnames(repro_avg) <- repro_avg[1,]
repro_avg <- repro_avg[-1,]
row.names(repro_avg) <- NULL

repro_avg %>% 
  kbl(booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position")) 
```


# Table S4. Plot density model
Model-averaged coefficients of the buffelgrass plot density model calculated from five top models (response variable: year-to-year change in plot density [individuals per m^2^]). Marginal R^2^ = 0.360–0.389; conditional R^2^ = 0.423–0.458.
```{r Table S4}
bgden_avg <- read.table("bgden_avg.txt", sep = "\t")
colnames(bgden_avg) <- bgden_avg[1,]
bgden_avg <- bgden_avg[-1,]
row.names(bgden_avg) <- NULL

bgden_avg %>% 
  kbl(booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position")) 
```


\newpage
# Table S5. Plot cover model
Model-averaged coefficients of the buffelgrass plot cover model calculated from six top models (response variable: year-to-year-change in plot cover [%]). Marginal R^2^ = 0.243–0.246; conditional R^2^ = 0.284–0.289.
```{r Table S5}
bgcov_avg <- read.table("bgcov_avg.txt", sep = "\t")
colnames(bgcov_avg) <- bgcov_avg[1,]
bgcov_avg <- bgcov_avg[-1,]
row.names(bgcov_avg) <- NULL

bgcov_avg %>% 
  kbl(booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position")) 
```


# Table S6. Seedling survival model
Model-averaged coefficients of the buffelgrass seedling survival model calculated from 15 top models (response variable: seedling survival [%]). Marginal R^2^ = 0.666–0.695; conditional R^2^ = 0.779–0.800.
```{r Table S6}
survival_avg <- read.table("survival_avg.txt", sep = "\t")
colnames(survival_avg) <- survival_avg[1,]
survival_avg <- survival_avg[-1,]
row.names(survival_avg) <- NULL

survival_avg %>% 
  kbl(booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "hold_position")) 
```

